---
title: "R Notebook"
output: html_notebook
---

# Setup
## libraries
```{r setup1, echo=TRUE}
library(tidyverse)
library(SqlRender)
library(DatabaseConnector)

connectionDetails <- createConnectionDetails(dbms="postgresql", 
                                             user="primed", password="primed", 
                                             server="localhost/primed", port=5432, 
                                             pathToDriver="./JdbcDrivers/")
```

Output directory for the results: medication_protocol

```{r setupdir}
dir.create("medication_protocol",showWarnings = FALSE)
```


## functions

```{r setup2, echo=TRUE}
find_drugs_by_ATC_class_concept_id <- function(class_concept_id) {
  connectionDetails <- createConnectionDetails(dbms="postgresql", 
                                             user="primed", password="primed", 
                                             server="localhost/primed", port=5432, 
                                             pathToDriver="./JdbcDrivers/")
  
  sql <- render("SELECT  d.concept_code, d.concept_name, 
        d.domain_id, d.concept_class_id, d.standard_concept,
        d.vocabulary_id,
        v.vocabulary_name
 FROM primed.concept_ancestor ca
        JOIN primed.concept d on d.concept_id = ca.descendant_concept_id
        JOIN primed.concept a on a.concept_id = ca.ancestor_concept_id
        JOIN primed.vocabulary v on d.vocabulary_id = v.vocabulary_id
 WHERE  ca.ancestor_concept_id = @class_concept_id
   AND  a.vocabulary_id = 'ATC'
   AND  d.domain_id = 'Drug'
   AND  (getdate() >= d.valid_start_date)
   AND  (getdate() <= d.valid_end_date);",
                    class_concept_id=class_concept_id)
  #print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  results <- querySql(conn,sql)
  disconnect(conn)
  return(results)
}


brand_name_search <- function(x) {
  print(paste("Searching for brand names of",x["CONCEPT_NAME"], " with ATC code ", x["CONCEPT_CODE"]," and OMOP Concept ID ",x["CONCEPT_ID"]))
  to_return <- find_drugs_by_ATC_class_concept_id(x["CONCEPT_ID"]) %>% 
           filter(grepl("Brand",CONCEPT_CLASS_ID)) %>% #& VOCABULARY_ID=="RxNorm") %>% 
           mutate(brand_name = str_extract(CONCEPT_NAME, "\\[.*\\]")) %>% 
           mutate(brand_name = str_replace_all(brand_name,"\\[","")) %>% 
           mutate(brand_name = str_replace_all(brand_name,"\\]","")) %>%
           mutate(brand_name = str_to_lower(brand_name))
  for(vocab in unique(to_return$VOCABULARY_ID)) {
    print(paste("Found ",length(unique(to_return[to_return$VOCABULARY_ID==vocab,]$brand_name))," brand names in ",vocab,":", paste(unique(to_return[to_return$VOCABULARY_ID==vocab,]$brand_name),collapse=", ")))
  }
  return(to_return)
}
```



# Sources

In the current analysis, we will use the following sources:

```{r sources1, echo=TRUE}
  sql <- render("SELECT  c.vocabulary_id, c.concept_class_id, count(*) as n
 FROM primed.concept c
 WHERE  c.domain_id = 'Drug'
 GROUP BY c.vocabulary_id, c.concept_class_id;")
  #print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  results <- querySql(conn,sql)
  disconnect(conn)
  print(results)
```

# Protocol

A search should begin with the chemical substance ([ATC 5th Level classification](https://www.who.int/tools/atc-ddd-toolkit/atc-classification)) of the drug. The chemical substance is the active ingredient of the drug.

## Metformin:

Chemical substance: "metformin"

```{r protocol1, echo=TRUE}
sql <- render("SELECT *
 FROM primed.concept c
 WHERE  
 c.vocabulary_id = 'ATC' AND c.concept_class_id = 'ATC 5th' AND c.concept_name LIKE '%metformin%'")
  #print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  metformin_concept_search_results <- querySql(conn,sql)
print(metformin_concept_search_results)

write_csv(metformin_concept_search_results,"medication_protocol/metformin_concept_search_results.csv")
```

### get brand names
```{r echo=TRUE, message=FALSE, warning=FALSE}
metformin_OMOP_CDM_brand_name_search <- bind_rows( apply(metformin_concept_search_results,1,brand_name_search))
data_frame(metformin=unique(metformin_OMOP_CDM_brand_name_search$brand_name))

write_csv(metformin_OMOP_CDM_brand_name_search,"medication_protocol/metformin_OMOP_CDM_brand_name_search.csv")

```

Check list against that from Drugbank.com. Which of the following were obtained with the OMOP CDM search and the RxNorm database?
```{r}

drug_list_string <- "Actoplus Met,Apo-Metformin,Avandamet,Fortamet,Glucophage,Glucovance,Glumetza,Glycon,Invokamet,Janumet,Jentadueto,Kazano,Kombiglyze,Komboglyze,Qternmet,Riomet,Segluromet,Synjardy,Trijardy,Velmetia,Xigduo,Zituvimet"
drug_list <- str_split(drug_list_string,",")[[1]]

table(str_to_lower(drug_list) %in% unique(metformin_OMOP_CDM_brand_name_search$brand_name))
str_to_lower(drug_list)[!str_to_lower(drug_list) %in% unique(metformin_OMOP_CDM_brand_name_search$brand_name)]

```


## Insulin degludec

Chemical substance: "insulin degludec"
### Find ATC codes
```{r echo=TRUE}
sql <- render("SELECT *
 FROM primed.concept c
 WHERE  
 c.vocabulary_id = 'ATC' AND c.concept_class_id = 'ATC 5th' AND c.concept_name LIKE '%insulin degludec%'")
  #print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  degludec_concept_search_results <- querySql(conn,sql)
print(degludec_concept_search_results)

write_csv(degludec_concept_search_results,"medication_protocol/degludec_concept_search_results.csv")
```
## get brand names

```{r}
degludec_OMOP_CDM_brand_name_search <- bind_rows( apply(degludec_concept_search_results,1,brand_name_search))
data_frame(degludec=unique(degludec_OMOP_CDM_brand_name_search$brand_name))

write_csv(degludec_OMOP_CDM_brand_name_search,"medication_protocol/degludec_OMOP_CDM_brand_name_search.csv")

```
```





