---
title: "Codes for vocabularies"
output:
  html_document: 
    df_print: kable
    toc: true
    number_sections: true
---

# NHGRI's PRIMED Consortium "EMR Codes for Legacy Project"

Release notes:
 - 15October2024: Added the codes to use for Trait queries
 - 30Sept2024.tsv: Added the codes to use for Drug queries

Table of Contents

```{r setup, include=FALSE}
library(DBI)
library(RPostgreSQL)
library(DatabaseConnector)
library(SqlRender)
```

## Common Data Model Vocabularies

Vocabularies were obtained through the [Athena Vocabulary Download API](https://athena.ohdsi.org/vocabulary/list). The latest update date is provided for each vocabulary. The AKM_30Sept2024 download list includes only the vocabularies selected by default.

If additional vocabularies are needed, please email Alisa Manning at amanning\@broadinstitute.org.

Detailed documentation about [Standardized Vocabularies](https://ohdsi.github.io/TheBookOfOhdsi/StandardizedVocabularies.html) are available in [The Book Of OHDSI](https://ohdsi.github.io/TheBookOfOhdsi/).

For documentation detailing the setup of a local database, please see [Set_up_database.html](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Set_up_database.html)

| ID (CDM v4.5) | CODE (CDM v5) | Name | Required | Latest Update |
|:--------------|:--------------|:--------------|:--------------|:--------------|
| 1 | SNOMED | Systematic Nomenclature of Medicine - Clinical Terms (IHTSDO) |  | 31-Jan-2024 |
| 2 | ICD9CM | International Classification of Diseases, Ninth Revision, Clinical Modification, Volume 1 and 2 (NCHS) |  | 30-Sep-2014 |
| 3 | ICD9Proc | International Classification of Diseases, Ninth Revision, Clinical Modification, Volume 3 (NCHS) |  | 30-Sep-2014 |
| 4 | CPT4 | Current Procedural Terminology version 4 (AMA) | EULA required | 05-May-2024 |
| 5 | HCPCS | Healthcare Common Procedure Coding System (CMS) |  | 30-Jun-2024 |
| 6 | LOINC | Logical Observation Identifiers Names and Codes (Regenstrief Institute) |  | 26-Feb-2024 |
| 8 | RxNorm | RxNorm (NLM) |  | 05-May-2024 |
| 9 | NDC | National Drug Code (FDA and manufacturers) |  | 24-Aug-2024 |
| 12 | Gender | OMOP Gender |  |  |
| 13 | Race | Race and Ethnicity Code Set (USBC) |  |  |
| 14 | CMS Place of Service | Place of Service Codes for Professional Claims (CMS) |  | 04-Aug-2024 |
| 21 | ATC | WHO Anatomic Therapeutic Chemical Classification |  | 02-Jul-2024 |
| 43 | Revenue Code | UB04/CMS1450 Revenue Codes (CMS) |  |  |
| 44 | Ethnicity | OMOP Ethnicity |  |  |
| 47 | NUCC | National Uniform Claim Committee Health Care Provider Taxonomy Code Set (NUCC) |  | 25-Jun-2018 |
| 48 | Medicare Specialty | Medicare provider/supplier specialty codes (CMS) |  | 25-Jun-2018 |
| 50 | SPL | Structured Product Labeling (FDA) |  | 24-Aug-2024 |
| 65 | Currency | International Currency Symbol (ISO 4217) |  |  |
| 70 | ICD10CM | International Classification of Diseases, Tenth Revision, Clinical Modification (NCHS) |  | 30-Sep-2024 |
| 71 | ABMS | Provider Specialty (American Board of Medical Specialties) |  | 25-Jun-2018 |
| 82 | RxNorm Extension | OMOP RxNorm Extension |  | 30-Jun-2024 |
| 128 | OMOP Extension | OMOP Extension (OHDSI) |  | 15-Jul-2024 |

Common Data Model Queries

The queries we used were based on the ODSI Query Library

The query used to generate the codes for traits, conditions and procedures is based on "**G06: Find concepts and their descendants that are covered by a given source code**

```{r eval=FALSE, echo=FALSE}

find_descendants_for_a_given_concept_id <- function(concept_id) {
  connectionDetails <- createConnectionDetails(dbms="postgresql", 
                                             user="primed", password="primed", 
                                             server="localhost/primed", port=5432, 
                                             pathToDriver="./JdbcDrivers/")
  
  sql <- render("SELECT
  c.concept_id               AS descendant_concept_id,
  c.concept_name             AS descendant_concept_name,
  c.concept_code             AS descendant_concept_code,
  c.concept_class_id         AS descendant_concept_class_id,
  c.vocabulary_id            AS vocabulary_id,
  a.min_levels_of_separation AS min_separation,
  a.max_levels_of_separation AS max_separation,
  c.invalid_reason           AS invalid_reason
FROM primed.concept_ancestor AS a
  JOIN primed.concept AS c ON a.descendant_concept_id = c.concept_id
WHERE 
  a.ancestor_concept_id != a.descendant_concept_id AND 
  a.ancestor_concept_id = @concept_id
ORDER BY vocabulary_id, min_separation
;",
                    concept_id=concept_id)
  print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  results <- querySql(conn,sql)
  disconnect(conn)
  return(results)
}

# G06: Find concepts and their descendants that are covered by a given source code
```

```{r eval=TRUE, echo=TRUE}

find_descendants_for_a_given_concept_id <- function(concept_id) {
  connectionDetails <- createConnectionDetails(dbms="postgresql", 
                                             user="primed", password="primed", 
                                             server="localhost/primed", port=5432, 
                                             pathToDriver="./JdbcDrivers/")
  
  sql <- render("WITH dm AS ( -- collect direct maps
    SELECT
      c1.concept_code     AS source_code,
      c1.vocabulary_id    AS source_vocabulary_id,
      c1.domain_id        AS source_domain_id,
      c2.concept_id       AS target_concept_id,
      c2.concept_name     AS target_concept_name,
      c2.concept_code     AS target_concept_code,
      c2.concept_class_id AS target_concept_class_id,
      c2.vocabulary_id    AS target_concept_vocab_id
    FROM primed.concept_relationship AS cr
      JOIN primed.concept AS c1 ON cr.concept_id_1 = c1.concept_id
      JOIN primed.concept AS c2 ON cr.concept_id_2 = c2.concept_id
    WHERE
      cr.relationship_id = 'Maps to' AND
      c1.concept_id IN (@concept_id) AND
      c1.vocabulary_id = 'SNOMED'
)
SELECT
  dm.source_code,
  dm.source_vocabulary_id,
  dm.source_domain_id,
  dc.concept_id       AS target_concept_id,
  dc.concept_name     AS target_concept_name,
  dc.concept_code     AS target_concept_code,
  dc.concept_class_id AS target_concept_class_id,
  dc.vocabulary_id    AS target_concept_vocab_id
FROM primed.concept_ancestor AS ca -- collect descendants which includes ancestor itself
  JOIN dm ON ca.ancestor_concept_id = dm.target_concept_id
  JOIN primed.concept AS dc ON ca.descendant_concept_id = dc.concept_id
WHERE dc.standard_concept = 'S'
;",
                    concept_id=concept_id)
#  print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  results <- querySql(conn,sql)
  disconnect(conn)
  return(results)
}


```

## Codes to use for Trait Queries

### White Blood Cell Count
```{r echo=TRUE}
white_blood_cell_count_OMOP_4298431 <- find_descendants_for_a_given_concept_id(4298431)
write.table(white_blood_cell_count_OMOP_4298431,file="white_blood_cell_count_OMOP_4298431_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(white_blood_cell_count_OMOP_4298431$TARGET_CONCEPT_CLASS_ID, white_blood_cell_count_OMOP_4298431$TARGET_CONCEPT_VOCAB_ID)
```
Download file: [white_blood_cell_count_OMOP_4298431_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/white_blood_cell_count_OMOP_4298431_AKM_15Oct2024.tsv)

### HbA1C
```{r echo=TRUE}
hba1c_OMOP_4184637 <- find_descendants_for_a_given_concept_id(4184637)
write.table(hba1c_OMOP_4184637,file="hba1c_OMOP_4184637_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(hba1c_OMOP_4184637$TARGET_CONCEPT_CLASS_ID, hba1c_OMOP_4184637$TARGET_CONCEPT_VOCAB_ID)
```
Download file: [hba1c_OMOP_4184637_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hba1c_OMOP_4184637_AKM_15Oct2024.tsv)

### SBP

```{r echo=TRUE}
sbp_OMOP_4152194 <- find_descendants_for_a_given_concept_id(4152194)
write.table(sbp_OMOP_4152194,file="sbp_OMOP_4152194_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(sbp_OMOP_4152194$TARGET_CONCEPT_CLASS_ID, sbp_OMOP_4152194$TARGET_CONCEPT_VOCAB_ID)
```

Download file: [sbp_OMOP_4152194_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/sbp_OMOP_4152194_AKM_15Oct2024.tsv)

```{r echo=TRUE}
dbp_OMOP_4154790 <- find_descendants_for_a_given_concept_id(4154790)
write.table(dbp_OMOP_4154790,file="dbp_OMOP_4154790_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(dbp_OMOP_4154790$TARGET_CONCEPT_CLASS_ID, dbp_OMOP_4154790$TARGET_CONCEPT_VOCAB_ID)
```

Download file: [dbp_OMOP_4154790_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/dbp_OMOP_4154790_AKM_15Oct2024.tsv)

### HDLC

```{r echo=TRUE}
hdlc_OMOP_4101713 <- find_descendants_for_a_given_concept_id(4101713)
write.table(hdlc_OMOP_4101713,file="hdlc_OMOP_4101713_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(hdlc_OMOP_4101713$TARGET_CONCEPT_CLASS_ID, hdlc_OMOP_4101713$TARGET_CONCEPT_VOCAB_ID)
```
Download file: [hdlc_OMOP_4101713_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hdlc_OMOP_4101713_AKM_15Oct2024.tsv)

### LDLC
```{r echo=TRUE}
ldlc_OMOP_4012479 <- find_descendants_for_a_given_concept_id(4012479)
write.table(ldlc_OMOP_4012479,file="ldlc_OMOP_4012479_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(ldlc_OMOP_4012479$TARGET_CONCEPT_CLASS_ID, ldlc_OMOP_4012479$TARGET_CONCEPT_VOCAB_ID)
```
Download file: [ldlc_OMOP_4012479_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/ldlc_OMOP_4012479_AKM_15Oct2024.tsv)

### TG
```{r echo=TRUE}
tg_OMOP_4017787 <- find_descendants_for_a_given_concept_id(4017787)
write.table(tg_OMOP_4017787,file="tg_OMOP_4017787_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(tg_OMOP_4017787$TARGET_CONCEPT_CLASS_ID, tg_OMOP_4017787$TARGET_CONCEPT_VOCAB_ID)
```

Download file: [tg_OMOP_4017787_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/tg_OMOP_4017787_AKM_15Oct2024.tsv)

### Weight
```{r echo=TRUE}

weight_OMOP_4099154 <- find_descendants_for_a_given_concept_id(4099154)
write.table(weight_OMOP_4099154,file="weight_OMOP_4099154_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(weight_OMOP_4099154$TARGET_CONCEPT_CLASS_ID, weight_OMOP_4099154$TARGET_CONCEPT_VOCAB_ID)
```

Download file: [weight_OMOP_4099154_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/weight_OMOP_4099154_AKM_15Oct2024.tsv)

### height
```{r echo=TRUE}
height_OMOP_607590 <- find_descendants_for_a_given_concept_id(607590)
write.table(height_OMOP_607590,file="height_OMOP_607590_AKM_15Oct2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)

table(height_OMOP_607590$TARGET_CONCEPT_CLASS_ID, height_OMOP_607590$TARGET_CONCEPT_VOCAB_ID)
```

Download file: [height_OMOP_607590_AKM_15Oct2024.tsv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/height_OMOP_607590_AKM_15Oct2024.tsv)

## Codes to use for Drugs Queries

This query is based on "D11: Find source codes by drug class"

Query source: <https://data.ohdsi.org/QueryLibrary/>

```{r echo=TRUE}
find_drugs_by_ATC_class_concept_id <- function(class_concept_id) {
  connectionDetails <- createConnectionDetails(dbms="postgresql", 
                                             user="primed", password="primed", 
                                             server="localhost/primed", port=5432, 
                                             pathToDriver="./JdbcDrivers/")
  
  sql <- render("SELECT  d.concept_code, d.concept_name, 
        d.domain_id, d.concept_class_id, d.standard_concept,
        d.vocabulary_id,
        v.vocabulary_name
 FROM primed.concept_ancestor ca
        JOIN primed.concept d on d.concept_id = ca.descendant_concept_id
        JOIN primed.concept a on a.concept_id = ca.ancestor_concept_id
        JOIN primed.vocabulary v on d.vocabulary_id = v.vocabulary_id
 WHERE  ca.ancestor_concept_id = @class_concept_id
   AND  a.vocabulary_id = 'ATC'
   AND  d.domain_id = 'Drug'
   AND  (getdate() >= d.valid_start_date)
   AND  (getdate() <= d.valid_end_date);",
                    class_concept_id=class_concept_id)
  #print(sql)
  sql <- translate(sql,targetDialect = "postgresql")
  conn <- connect(connectionDetails)
  results <- querySql(conn,sql)
  disconnect(conn)
  return(results)
}

```

### Diabetes Medications

ATC 2nd level A10: DRUGS USED IN DIABETES <https://athena.ohdsi.org/search-terms/terms/21600712>

```{r echo=TRUE}
drugs_used_in_diabetes_OMOP_21600712 <- find_drugs_by_ATC_class_concept_id(21600712)
write.table(drugs_used_in_diabetes_OMOP_21600712,file="drugs_used_in_diabetes_OMOP_21600712_AKM_30Sept2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r echo=TRUE}
table(drugs_used_in_diabetes_OMOP_21600712$CONCEPT_CLASS_ID,drugs_used_in_diabetes_OMOP_21600712$VOCABULARY_ID)
```

#### Download Current Release and Search/Filter Table

[Drugs used in Diabetes](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Drugs-used-in-Diabetes.html) for data download.

### Blood Pressure Medications

ATC 2nd level C02: ANTIHYPERTENSIVES <https://athena.ohdsi.org/search-terms/terms/21600381>

```{r echo=TRUE}
drugs_used_in_hypertension_OMOP_21600381 <- find_drugs_by_ATC_class_concept_id(21600381)
write.table(drugs_used_in_hypertension_OMOP_21600381,file="drugs_used_in_hypertension_OMOP_21600381_AKM_30Sept2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r  echo=TRUE}
table(drugs_used_in_hypertension_OMOP_21600381$CONCEPT_CLASS_ID, drugs_used_in_hypertension_OMOP_21600381$VOCABULARY_ID)
```

#### Download Current Release and Search/Filter Table

See [Blood Pressure Lowering Medications](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Blood-Pressure-Lowering-Medications.html) for data download.

### Lipid Lowering Medication

ATC 2nd level C10: LIPID MODYFYING AGENTS <https://athena.ohdsi.org/search-terms/terms/21601853>

```{r echo=TRUE}
drugs_used_in_lipid_OMOP_21601853 <- find_drugs_by_ATC_class_concept_id(21601853)
write.table(drugs_used_in_lipid_OMOP_21601853,file="drugs_used_in_lipid_OMOP_21601853_AKM_30Sept2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r  echo=TRUE}
table(drugs_used_in_lipid_OMOP_21601853$CONCEPT_CLASS_ID,
      drugs_used_in_lipid_OMOP_21601853$VOCABULARY_ID)
```

#### Download Current Release and Search/Filter Table

See [Lipid Lowering Medications](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Lipid-Lowering-Medications.html) for data download.
