---
title: "PRIMED Consortium Legacy Project - Codes for Biobanks"
output:
  html_document: 
    df_print: kable
    toc: true
    number_sections: true
---
```{r}
# defaults for every cell
# do not show messages or warnings

```


# PRIMED Consortium Legacy Project - Codes for Biobanks

Release notes:
-   27 November 2024: Added "brand name medication" query and code files which provide brand name medications for every 5th level ATC code
-   18 November 2024: Refactored Exclusion codes to include additional vocabularies and simplify output
-   22 October 2024: Uploaded codes for Blood Cancers, HIV, End Stage Kidney Disease, spleen disorders, Fibrosis, Cirrhosis of the Liver
-   21 October2024: Uploaded Codes for Blood Disorders
-   17 October2024: Fixed the codes to use for Trait queries. Started on Conditions
-   15October2024: Added the codes to use for Trait queries
-   30Sept2024.tsv: Added the codes to use for Drug queries


```{r setup, include=FALSE}
library(DBI)
library(RPostgreSQL)
library(DatabaseConnector)
library(SqlRender)
library(dplyr)
library(readr)
```

## Common Data Model Vocabularies

Vocabularies were obtained through the [Athena Vocabulary Download API](https://athena.ohdsi.org/vocabulary/list). The latest update date is provided for each vocabulary. The AKM_30Sept2024 download list includes only the vocabularies selected by default.

If additional vocabularies are needed, please email Alisa Manning at amanning\@broadinstitute.org.

Detailed documentation about [Standardized Vocabularies](https://ohdsi.github.io/TheBookOfOhdsi/StandardizedVocabularies.html) are available in [The Book Of OHDSI](https://ohdsi.github.io/TheBookOfOhdsi/).

For documentation detailing the setup of a local database, please see [Set_up_database.html](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Set_up_database.html)

```{r}
in.vocabulary <- read_delim("vocabulary_download_v5_{d3fc0e41-71fa-4f84-9576-80ed6651d6a5}_1731967602015/VOCABULARY.csv", col_types = cols(
  vocabulary_id = col_character(),
  vocabulary_name = col_character(),
  vocabulary_reference = col_character(),
  vocabulary_version = col_character(),
  vocabulary_concept_id = col_integer()))

in.vocabulary
```

Common Data Model Queries

The queries we used were based on the ODSI Query Library

The query used to generate the codes for traits, conditions and procedures is based on "**G06: Find concepts and their descendants that are covered by a given source code**


```{r eval=TRUE, echo=TRUE}
source("find_descendants_and_mappings_for_a_given_concept_id.R")
```

## Codes to use for Trait Queries

### White Blood Cell Count

```{r echo=TRUE}
white_blood_cell_count_OMOP_4298431 <- find_descendants_and_mappings_for_a_given_concept_id(4298431, file_label="white_blood_cell_count_SNOMED_4298431_AKM_17Oct2024")

```

Download files:

-   [white_blood_cell_count_SNOMED_4298431_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/white_blood_cell_count_SNOMED_4298431_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [white_blood_cell_count_SNOMED_4298431_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/white_blood_cell_count_SNOMED_4298431_AKM_17Oct2024_UNIQUE_CODES.csv)

### Leukocytes \| Blood \| Hematology and Cell counts

```{r}
leukocytes_in_blood_LOINC_37043992 <- find_descendants_and_mappings_for_a_given_concept_id(37043992, file_label="leukocytes_in_blood_LOINC_37043992_AKM_17Oct2024")
```

Download files:

-   [leukocytes_in_blood_LOINC_37043992_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/leukocytes_in_blood_LOINC_37043992_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [leukocytes_in_blood_LOINC_37043992_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/leukocytes_in_blood_LOINC_37043992_AKM_17Oct2024_UNIQUE_CODES.csv)

### HbA1C

```{r echo=TRUE}
hba1c_SNOMED_4184637 <- find_descendants_and_mappings_for_a_given_concept_id(4184637, file_label="hba1c_SNOMED_4184637_AKM_17Oct2024")

```

Download file:

-   [hba1c_SNOMED_4184637_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hba1c_SNOMED_4184637_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [hba1c_SNOMED_4184637_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hba1c_SNOMED_4184637_AKM_17Oct2024_UNIQUE_CODES.csv)

```{r}
hba1c_LOINC_1621295 <- find_descendants_and_mappings_for_a_given_concept_id(1621295, file_label="hba1c_LOINC_1621295_AKM_17Oct2024")
```

Download files: 

- [hba1c_LOINC_1621295_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hba1c_LOINC_1621295_AKM_17Oct2024_ALL_MAPPINGS.csv) 
- [hba1c_LOINC_1621295_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hba1c_LOINC_1621295_AKM_17Oct2024_UNIQUE_CODES.csv)

### SBP

```{r echo=TRUE}
sbp_SNOMED_4152194 <- find_descendants_and_mappings_for_a_given_concept_id(4152194, file_label="sbp_SNOMED_4152194_AKM_17Oct2024")
```

Download file:

-   [sbp_SNOMED_4152194_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/sbp_SNOMED_4152194_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [sbp_SNOMED_4152194_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/sbp_SNOMED_4152194_AKM_17Oct2024_UNIQUE_CODES.csv)

```{r}
sbp_LOINC_3004249 <- find_descendants_and_mappings_for_a_given_concept_id(3004249, file_label="sbp_LOINC_3004249_AKM_17Oct2024")
```

Download file:

-   [sbp_LOINC_3004249_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/sbp_LOINC_3004249_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [sbp_LOINC_3004249_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/sbp_LOINC_3004249_AKM_17Oct2024_UNIQUE_CODES.csv)

```{r echo=TRUE}
dbp_SNOMED_4154790 <- find_descendants_and_mappings_for_a_given_concept_id(4154790, file_label="dbp_SNOMED_4154790_AKM_17Oct2024")

```

Download file:

-   [dbp_SNOMED_4154790_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/dbp_SNOMED_4154790_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [dbp_SNOMED_4154790_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/dbp_SNOMED_4154790_AKM_17Oct2024_UNIQUE_CODES.csv)

```{r}
dbp_LOINC_3012888 <- find_descendants_and_mappings_for_a_given_concept_id(3012888, file_label="dbp_LOINC_3012888_AKM_17Oct2024")
```

Download file:

-   [dbp_LOINC_3012888_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/dbp_LOINC_3012888_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [dbp_LOINC_3012888_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/dbp_LOINC_3012888_AKM_17Oct2024_UNIQUE_CODES.csv)

### HDLC

```{r echo=TRUE}
hdlc_SNOMED_4101713 <- find_descendants_and_mappings_for_a_given_concept_id(4101713, file_label="hdlc_SNOMED_4101713_AKM_17Oct2024")
```

Download file:

-   [hdlc_SNOMED_4101713_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hdlc_SNOMED_4101713_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [hdlc_SNOMED_4101713_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/hdlc_SNOMED_4101713_AKM_17Oct2024_UNIQUE_CODES.csv)

### LDLC

```{r echo=TRUE}
ldlc_SNOMED_4012479 <- find_descendants_and_mappings_for_a_given_concept_id(4012479, file_label="ldlc_SNOMED_4012479_AKM_17Oct2024")
```

Download file:

-   [ldlc_SNOMED_4012479_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/ldlc_SNOMED_4012479_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [ldlc_SNOMED_4012479_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/ldlc_SNOMED_4012479_AKM_17Oct2024_UNIQUE_CODES.csv)

### TG

```{r echo=TRUE}
tg_SNOMED_4032789 <- find_descendants_and_mappings_for_a_given_concept_id(4032789, file_label="tg_SNOMED_4032789_AKM_17Oct2024")
```

Download file:

-   [tg_SNOMED_4032789_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/tg_SNOMED_4032789_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [tg_SNOMED_4032789_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/tg_SNOMED_4032789_AKM_17Oct2024_UNIQUE_CODES.csv)

### Weight

```{r echo=TRUE}

weight_SNOMED_4099154 <- find_descendants_and_mappings_for_a_given_concept_id(4099154, file_label="weight_SNOMED_4099154_AKM_17Oct2024")
```

Download file:

-   [weight_SNOMED_4099154_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/weight_SNOMED_4099154_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [weight_SNOMED_4099154_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/weight_SNOMED_4099154_AKM_17Oct2024_UNIQUE_CODES.csv)

### height

```{r echo=TRUE}
height_SNOMED_607590 <- find_descendants_and_mappings_for_a_given_concept_id(607590, file_label="height_SNOMED_607590_AKM_17Oct2024")
```

Download file:

-   [height_SNOMED_607590_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/height_SNOMED_607590_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [height_SNOMED_607590_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/height_SNOMED_607590_AKM_17Oct2024_UNIQUE_CODES.csv)

## Codes to use for Condition Exclusions

### Codes for Blood Disorders

SNOMED: Disorder of cellular component of blood
SNOMED CODE: 414022008
OMOP CONCEPT ID: 443723
SNOMED: Disorder of hematopoietic structure 
SNOMED CODE: 414027002
OMOP CONCEPT ID: 317248
SNOMED: Agranulocytosis
SNOMED CODE: 17182001
OMOP CONCEPT ID: 440689
SNOMED: Bone marrow transplant present
SNOMED CODE: 737300001
OMOP CONCEPT ID: 42537745

```{r}
blood_disorders_SNOMED <- find_descendants_and_mappings_for_a_given_concept_id(c(443723,317248,440689,42537745), file_label="blood_disorders_SNOMED_AKM_21October2024")
```

### Codes for Blood Cancers

SNOMED: Leukemia
SNOMED CODE: 93143009
OMOP CONCEPT ID: 317510
SNOMED: Malignant lymphoma
SNOMED CODE: 118600007
OMOP CONCEPT ID: 432571
SNOMED: Disorder of hematopoietic morphology
SNOMED CODE: 414026006
OMOP CONCEPT ID: 4185470

```{r}
blood_cancers_SNOMED <- find_descendants_and_mappings_for_a_given_concept_id(c(317510,432571,4185470), file_label="blood_cancers_SNOMED_AKM_18November2024")
```

### Codes for HIV infection

```{r}
hiv_infection_SNOMED_439727 <- find_descendants_and_mappings_for_a_given_concept_id(c(439727), file_label="hiv_infection_SNOMED_439727_AKM_27November2024")
```

### Codes for End Stage Kidney Disease

SNOMED: Dependence on renal dialysis
SNOMED Code: 105502003
OMOP Concept ID: 4019967

SNOMED: End-stage renal disease
SNOMED Code: 46177005
OMOP Concept ID: 193782

SNOMED: Chronic kidney disease stage 5
SNOMED Code: 433146000
OMOP Concept ID: 443611

SNOMED Procedure: Renal dialysis
SNOMED Code: 265764009
OMOP Concept ID: 4146536


```{r}
ESKD_SNOMED <- find_descendants_and_mappings_for_a_given_concept_id(c(4019967, 193782,443611,4146536), file_label="ESKD_SNOMED_AKM_27Nov2024")

```

#### End stage renal disease

```{r}
end_stage_renal_disease_SNOMED_4019967 <- find_descendants_and_mappings_for_a_given_concept_id(4019967, file_label="end_stage_renal_disease_SNOMED_4019967_AKM_17Oct2024")
```

Files to download:

-   [end_stage_renal_disease_SNOMED_4019967_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/end_stage_renal_disease_SNOMED_4019967_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [end_stage_renal_disease_SNOMED_4019967_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/end_stage_renal_disease_SNOMED_4019967_AKM_17Oct2024_UNIQUE_CODES.csv)

#### Chronic kidney disease stage 5

```{r}
chronic_kidney_disease_stage_5_SNOMED_443611 <- find_descendants_and_mappings_for_a_given_concept_id(443611, file_label="chronic_kidney_disease_stage_5_SNOMED_443611_AKM_17Oct2024")
```

Download Files:

-   [chronic_kidney_disease_stage_5_SNOMED_443611_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/chronic_kidney_disease_stage_5_SNOMED_443611_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [chronic_kidney_disease_stage_5_SNOMED_443611_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/chronic_kidney_disease_stage_5_SNOMED_443611_AKM_17Oct2024_UNIQUE_CODES.csv)

#### Renal dialysis

```{r}
renal_dialysis_SNOMED_4146536 <- find_descendants_and_mappings_for_a_given_concept_id(4146536, file_label="renal_dialysis_SNOMED_4146536_AKM_17Oct2024")
```

Download Files:

-   [renal_dialysis_SNOMED_4146536_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/renal_dialysis_SNOMED_4146536_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [renal_dialysis_SNOMED_4146536_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/renal_dialysis_SNOMED_4146536_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes for Splenectomy and spleen disorders

Files with ICD-9 and ICD-10 codes

-   [ICD-9-XXXX.csv]()
-   [ICD-10-XXX.csv]()

#### Disorder of spleen

```{r}
disorder_of_spleen_SNOMED_201265 <- find_descendants_and_mappings_for_a_given_concept_id(201265, file_label="disorder_of_spleen_SNOMED_201265_AKM_17Oct2024")
```

Files to download:

-   [disorder_of_spleen_SNOMED_201265_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/disorder_of_spleen_SNOMED_201265_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [disorder_of_spleen_SNOMED_201265_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/disorder_of_spleen_SNOMED_201265_AKM_17Oct2024_UNIQUE_CODES.csv)

#### Splenectomy

```{r}
splenectomy_SNOMED_4125482 <- find_descendants_and_mappings_for_a_given_concept_id(4125482, file_label="splenectomy_SNOMED_4125482_AKM_17Oct2024")
```

Files to download:

-   [splenectomy_SNOMED_4125482_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/splenectomy_SNOMED_4125482_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [splenectomy_SNOMED_4125482_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/splenectomy_SNOMED_4125482_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes for Fibrosis, Cirrhosis of the Liver

Files with ICD-9 and ICD-10 codes

-   [ICD-9-XXXX.csv]()
-   [ICD-10-XXX.csv]()

#### Cirrhosis of liver

```{r}
cirrhosis_of_liver_SNOMED_4064161 <- find_descendants_and_mappings_for_a_given_concept_id(4064161, file_label="cirrhosis_of_liver_SNOMED_4064161_AKM_17Oct2024")
```

Files to download:

-   [cirrhosis_of_liver_SNOMED_4064161_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/cirrhosis_of_liver_SNOMED_4064161_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [cirrhosis_of_liver_SNOMED_4064161_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/cirrhosis_of_liver_SNOMED_4064161_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes for blood transfusions

#### Transfusion of blood product

```{r}
transfusion_of_blood_product_SNOMED_4024656 <- find_descendants_and_mappings_for_a_given_concept_id(4024656, file_label="transfusion_of_blood_product_SNOMED_4024656_AKM_17Oct2024")
```

-   [transfusion_of_blood_product_SNOMED_4024656_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/transfusion_of_blood_product_SNOMED_4024656_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [transfusion_of_blood_product_SNOMED_4024656_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/transfusion_of_blood_product_SNOMED_4024656_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes for Diabetes

#### Diabetes mellitus

```{r}
diabetes_mellitus_SNOMED_201820 <- find_descendants_and_mappings_for_a_given_concept_id(201820, file_label="diabetes_mellitus_SNOMED_201820_AKM_17Oct2024")
```

Files to download:

-   [diabetes_mellitus_SNOMED_201820_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/diabetes_mellitus_SNOMED_201820_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [diabetes_mellitus_SNOMED_201820_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/diabetes_mellitus_SNOMED_201820_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes to use for Cancer

#### Malignant neoplastic disease

```{r}
malignant_neoplastic_disease_SNOMED_443392 <- find_descendants_and_mappings_for_a_given_concept_id(443392, file_label = "malignant_neoplastic_disease_SNOMED_443392_AKM_17Oct2024")
```

Files to download:

-   [malignant_neoplastic_disease_SNOMED_443392_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/malignant_neoplastic_disease_SNOMED_443392_AKM_17Oct2024_ALL_MAPPINGS.csv)
-   [malignant_neoplastic_disease_SNOMED_443392_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/malignant_neoplastic_disease_SNOMED_443392_AKM_17Oct2024_UNIQUE_CODES.csv)

### Codes used for pregnancy

#### Pregnancy

```{r}
pregnancy_SNOMED_4299535 <- find_descendants_and_mappings_for_a_given_concept_id(4299535, file_label="pregnancy_SNOMED_4299535_AKM_17Oct2024")
```
Files to download:

-  [pregnancy_SNOMED_4299535_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/pregnancy_SNOMED_4299535_AKM_17Oct2024_ALL_MAPPINGS.csv)
- [pregnancy_SNOMED_4299535_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/pregnancy_SNOMED_4299535_AKM_17Oct2024_UNIQUE_CODES.csv)

#### Complication occurring during pregnancy

```{r}
complication_occurring_during_pregnancy_SNOMED_43530950 <- find_descendants_and_mappings_for_a_given_concept_id(43530950, file_label="complication_occurring_during_pregnancy_SNOMED_43530950_AKM_17Oct2024")
```

Files to download:

- [complication_occurring_during_pregnancy_SNOMED_43530950_AKM_17Oct2024_ALL_MAPPINGS.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/complication_occurring_during_pregnancy_SNOMED_43530950_AKM_17Oct2024_ALL_MAPPINGS.csv)
- [complication_occurring_during_pregnancy_SNOMED_43530950_AKM_17Oct2024_UNIQUE_CODES.csv](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/complication_occurring_during_pregnancy_SNOMED_43530950_AKM_17Oct2024_UNIQUE_CODES.csv)

## Codes to use for Drugs Queries

This query is based on "D11: Find source codes by drug class"

Query source: <https://data.ohdsi.org/QueryLibrary/>

```{r echo=TRUE}
source("medication_functions.R")
# create directory called medication_codes
dir.create("medication_codes", showWarnings = FALSE)
```

### Diabetes Medications
#### ATC 2nd level: A10: DRUGS USED IN DIABETES
ATC 2nd level A10: DRUGS USED IN DIABETES <https://athena.ohdsi.org/search-terms/terms/21600712>

Download Current Release and Search/Filter Table: [Drugs used in Diabetes](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Drugs-used-in-Diabetes.html) for data download.

```{r echo=TRUE, warning=FALSE}
drugs_used_in_diabetes_OMOP_21600712 <- find_drugs_by_ATC_class_concept_id(21600712)
write.table(drugs_used_in_diabetes_OMOP_21600712,file="medication_codes/drugs_used_in_diabetes_OMOP_21600712_AKM_27November2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r echo=TRUE}
table(drugs_used_in_diabetes_OMOP_21600712$CONCEPT_CLASS_ID,drugs_used_in_diabetes_OMOP_21600712$VOCABULARY_ID)
```

#### Brand name search will all 5th level ATC codes
```{r echo=TRUE,warning=FALSE}
brand_names_drugs_used_in_diabetes_OMOP_21600712 <-
  bind_rows( apply(drugs_used_in_diabetes_OMOP_21600712 %>% filter(CONCEPT_CLASS_ID == "ATC 5th"),1,brand_name_search))

write.table(brand_names_drugs_used_in_diabetes_OMOP_21600712,file="medication_protocol/brand_names_drugs_used_in_diabetes_OMOP_21600712_AKM_27November2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)


```

#### Count of ATC code vs Brand Name

```{r}
table(brand_names_drugs_used_in_diabetes_OMOP_21600712$ATC_5th, brand_names_drugs_used_in_diabetes_OMOP_21600712$VOCABULARY_ID)
```

#### Count of ATC Codes vs CONCEPT_CLASS_ID
```{r}
table(brand_names_drugs_used_in_diabetes_OMOP_21600712$ATC_5th, brand_names_drugs_used_in_diabetes_OMOP_21600712$CONCEPT_CLASS_ID)
```

### Blood Pressure Medications

ATC 2nd level C02: ANTIHYPERTENSIVES <https://athena.ohdsi.org/search-terms/terms/21600381>

```{r echo=TRUE}
drugs_used_in_hypertension_OMOP_21600381 <- find_drugs_by_ATC_class_concept_id(21600381)
write.table(drugs_used_in_hypertension_OMOP_21600381,file="drugs_used_in_hypertension_OMOP_21600381_AKM_30Sept2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r  echo=TRUE}
table(drugs_used_in_hypertension_OMOP_21600381$CONCEPT_CLASS_ID, drugs_used_in_hypertension_OMOP_21600381$VOCABULARY_ID)
```

#### Download Current Release and Search/Filter Table

See [Blood Pressure Lowering Medications](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Blood-Pressure-Lowering-Medications.html) for data download.

### Lipid Lowering Medication

ATC 2nd level C10: LIPID MODYFYING AGENTS <https://athena.ohdsi.org/search-terms/terms/21601853>

```{r echo=TRUE}
drugs_used_in_lipid_OMOP_21601853 <- find_drugs_by_ATC_class_concept_id(21601853)
write.table(drugs_used_in_lipid_OMOP_21601853,file="drugs_used_in_lipid_OMOP_21601853_AKM_30Sept2024.tsv",sep="\t",quote=FALSE,row.names=FALSE)
```

#### Table of CONCEPT_CLASS_ID vs VOCABULARY_ID

```{r  echo=TRUE}
table(drugs_used_in_lipid_OMOP_21601853$CONCEPT_CLASS_ID,
      drugs_used_in_lipid_OMOP_21601853$VOCABULARY_ID)
```

#### Download Current Release and Search/Filter Table

See [Lipid Lowering Medications](https://manning-lab.github.io/primed_consortia_legacy_project_phenotypes/Lipid-Lowering-Medications.html) for data download.
